Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources

### 2019-01-23 paper.tex
Update problem section and graph sizes

\begin{document}
\title[Projecting Globel Filesystems into Local Containers.]{Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources}

@@

\begin{abstract}
Scientific applications often depend on large, complex software repositories containing libraries, tools, and frameworks in multiple versions and platforms to support many researchers.
As a practical necessity, software repos are most commonly distributed via shared or global filesystems.
In moving existing research to HPC environments, however,
scientists must work around restrictions that can interfere with global software delivery methods.
To illustrate these challenges,
we examine CVMFS, the CernVM File System.
CVMFS is a global, distributed filesystem that provides read-only access to experiment software and configurations for high energy physics research at CERN.
CVMFS requires a number of network and OS configurations that are incompatible with a restrictive HPC environment.
In this work, we introduce a set of tools and procedures for \emph{projecting} the global filesystem (or parts thereof) into local container images suitable for HPC sites.
As part of our evaluation,
we found that using local approximations of a global filesystem tends to introduce problems of its own in terms of compute and storage overhead.
We thus examined several declarative and profiling-based approaches to creating more fine-grained projections to fit application requirements,
and propose a strategy for efficiently managing these projected container images at an HPC site.
Using our projection tools and image management strategy,
we implemented a proof-of-concept submission system for running existing HEP analysis tasks in an HPC environment.
\end{abstract}

@@

@@

\section{Background and Challenges}

@@

\begin{table*}[t]
\begin{center}
\begin{tabular}{|c|r r r|r r r|}
\hline
CVMFS & & ls -lr & & & Shrinkwrap & \\
Repository & Total Size (MB) & inode Count & Time to walk & 
  Data Files & Deduplicated Data Files & Deduplicated Data (MB) \\ \hline
%alice-ocdb & 482,642,307,428 & 1,621,661 & 2:22 &
% 1,621,282 & 1,595 & 24,433,131,203\\
%alice & 3,980,875,747,986 & 27,322,350 & 51:53 & 
%  797,404 & 23,996,541 & 2,098,509,015,172\\
alice & 3,980,875 & 27,322,000 & 51:53 & 
  797,000 & 23,997,000 & 2,098,509\\
%atlas-condb & 719,548,941,240 & 11,781 & 0:04 &
%  10,940 & 229 & 471,492,086 \\
%atlas & 6,195,013,689,056 & 120,598,601 & 3:48:07 &
% 8,823,179 & 85,554,444 & 3,760,712,059,548 \\
atlas & 6,195,013 & 120,599,000 & 3:48:07 &
 8,823,000 & 85,554,000 & 3,760,712 \\
%boss & 120,048,527,315 & 2,418,025 & 4:18 &
% 256,061 & 1,868,192 & 72,301,955,666\\
boss & 120,048 & 2,418,000 & 4:18 &
 256,000 & 1,868,000 & 72,302\\
%cms & 581,789,406,847 & 26,224,567 & 1:07:30 &
% 18,667,178 & 101,448,867 & *2,242,658,825,701\\
cms & 581,789 & 26,225,000 & 1:07:30 &
 18,667,000 & 101,449,000 & *2,242,658\\
%geant4 & 297,571,283,648 & 7,128,024 & 10:36 &
% 323,280 & 6,219,891 & 68,381,969,660 \\
geant4 & 297,571 & 7,128,000 & 10:36 &
 323,000 & 6,220,000 & 68,382 \\
%grid & 30,642,826,259 & 714,235 & 1:17 &
% 141,622 & 583,923 & 18,804,003,040 \\
grid & 30,643 & 714,000 & 1:17 &
 142,000 & 584,000 & 18,804 \\
%lhcb & 2,134,893,079,440 & 33,736,802 & 1:01:26 &
% 10,032,339 & 20,406,235 & 368,070,277,634 \\
lhcb & 2,134,893 & 33,737,000 & 1:01:26 &
 10,032,000 & 20,406,000 & 368,070 \\
%na61 & 5,222,967,734 & 25,602 & 0:02 &
% 8,609 & 15,001 & 2,425,507,431 \\
na61 & 5,222 & 26,000 & 0:02 &
 9,000 & 15,000 & 2,426 \\
%sft & 3,996,527,275,237 & 128,608,792 & 4:17:58 &
% 6,821,699 & 63,210,139 & 2,016,609,810,573 \\ 
sft & 3,996,527 & 128,609,000 & 4:17:58 &
 6,822,000 & 63,210,000 & 2,016,610 \\ 
 \hline
\end{tabular}
\caption{CVMFS repository sizes.}
\label{tab:repo-sizes}
This table shows the size, number of inodes (rounded to nearest thousand), and the amount
of duplicate data that is eliminated using CVMFS.
\end{center}
\end{table*}

@@

\if 0
spread across everyone, lot of commonality/overlap
few with very high usage, very large number of rarely used
lot not observed
\fi

@@

Figure~\ref{fig:series} shows a single simulation of the image management system,
with several notable details.
First, we see that most of the operations are merges.
This makes sense, as this simulation was at fairly high $\alpha$.
The total bytes written also closely tracks merges,
indicating that merging is the dominant source of I/O.
We can still see inserts over the course of the simulation.
At more extreme $\alpha$ values,
we would expect to see one of these operations almost exclusively.
As the data in cache continues to rise,
we eventually see the cache limit,
after which the delete count starts increasing.
Over the course of the simulation,
inserts and deletes are filling and emptying the cache such that it remains close to its storage limit.
We also observe the number of cache hits continue to rise despite deletions.
As we will see, merging allows for a greater proportion of hits even if the amount of data remains constant.
This is because frequently used data can be merged,
reducing duplication.
The cache limit then ensures that infrequently used parts are eventually removed.

@@

\begin{figure*}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache_percent_plt.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request_percent_plt.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses_percent_plt.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses_percent_plt.pdf}
\label{fig:sensitivity}
\if 0
[t] "just buy more cache"
efficiency is worse, since cache holds on to lots of questionable data
loads of opportunities to try merges, so we see a lot of unnecessary merges
less force removing bad choices, so more cruft can accumulate in many images
better cache utilization with tiny cache, but of course more thrashing
at scale, bad idea
large cache is too pricey at scale
[m] varying reuse doesn't affect too much
below 100 doesn't seem full/steady state yet
cache isn't settled? so haven't dropped bad merges; more unique data in cache boosts efficiency
[b] high number of requests, repeats don't matter
unique data dominates
hard to tell difference between [m] and [b], so not too sensitive to these
once you hit steady state, jobs count/repeats has little effect on efficiency
general trends of merging, not interesting by themselves
\fi
\end{figure*}

@@

\section{Related Work}
TODO

>>>>

\begin{document}
\title[Projecting Global Filesystems into Local Containers.]{Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources}

@@

\begin{abstract}
Scientific applications often depend on large, complex software repositories containing libraries, tools, and frameworks in multiple versions and platforms to support many researchers.
Such repositories are distributed to hundreds of major computing
facilities around the world using a global filesystem like CVMFS.
However, not all HPC facilities permit execution nodes to
mount and access global network resources, instead encouraging
applications to make use of local filesystems and container
technologies.  In this work, we present a technique for efficiently.
\emph{projecting} a global filesystem into local resources.
This consists of creating minimal local containers that meet
the needs of specific jobs, balancing between the extremes
of having many small single-purpose containers and having one
large all-purpose containers.  We evaluate these techniques
using data from production instances of the CVMFS global filesystem
used by a wide variety of scientific collaborations.
\end{abstract}

@@

Global scientific experiment collaborations like ATLAS, CMS, LIGO,
and LSST make use of hundreds of thousands of cores spread across
many computing facilities around the world in order to achieve the massive scale needed simulate and analyze data from high bandwidth
scientific instruments.  In order to ensure consistency of software
running at many different sites, collaborations establish global
software repositories that publish the tested and trusted versions
of software to be used for production runs.  These repositories
can be quite large, consisting of thousands of software packages
consuming millions of files and terabytes of data.  Software
such as CVMFS~\cite{globalfs-cise-2015} is used to distribute
these complex software stacks around the world.

However, not all computing facilities make it easy to access global filesystems at runtime.
In particular, national HPC facilities that are not associated with any
particular experiment often prevent network connections between
individual worker nodes and the outside world, for reasons
of both security and system efficiency.  Such facilities instead
encourage the use of local high performance filesystems and container
technologies to deliver software to individual worker nodes.
To make use of such sites, some mapping between the global
filesystem and the locally available technologies must be provided.

In this paper, we describe the problem of {\bf projecting} a global
filesystem into the resources available at a local HPC facility.
There are two main components to this problem.

First, for a given job, we must have a means of identifying the
subset of the global filesystem that is relevant for the current
execution.  This can be done either by dynamic tracing of the resources
actually used, or by static analysis of its dependency tree.
We describe Shrinkwrap, a tool that performs this analysis on
a Unix application, and produces a minimal container
that can be used to execute the application, or one closely related.
We analyze the efficiency of this tool on a selection of high
energy physics applications, and demonstrate storage efficiency of this technique, and measure the performance and storage efficiency.

Second, for a large number of jobs, we must have a method for
managing the large number of containers that will be created.
If one small container is created for every single job,
site storage will overflow with duplicated data.  On the other
hand, if one large container is constructed to satisfy all
possible jobs, it will overflow constraints at the worker nodes.
To address this, we develop an online algorithm that manages
a shared storage space with multiple containers, choosing to
either create, delete, or merge containers as new requests arrive.
The key parameter ($\alpha$) to this algorithm allows for a
tradeoff between the efficiency of cache storage and container size.

Both techniques are evaluated in light of the properties
of large global filesystems used in production by high
energy physics applications.  In particular, we exploit the
property that many jobs share a small set of frequently-used
dependencies, and have a long tail of infrequently-used dependencies.
We found that our approaches can effectively manage the container images for a continuous stream of submissions with varying software dependencies and versions with less storage overhead than creating a new image per application.
In addition, our results show that applications can still take advantage of precise dependency and version specifications to reduce image creation and transfer overhead compared to simple mirroring.

@@

\begin{figure}
\includegraphics[width=\linewidth]{drawings/methods.pdf}
\label{choices}
\caption{Global Filesystem Access Methods}
\end{figure}

\section{Accessing Global Filesystems}

Modern scientific applications depend upon complex software
repositories that consist of terabytes of data and millions of files.
Unlike "traditional" high performance applications that are
single, statically-linked executables, modern applications are
agglomerations of software that consist of high level scripts,
multiple executables written in different languages, support
libraries that may be dynamically linked, and configuration and
data files loaded at runtime.

In a well-organized project, the necessary software dependencies
are collected together into a global repository that is accessible
from all of the sites that intend to use it.  A common arrangement
is to publish a filesystem that can be mounted from each execution node that intends to use it.  For example, the CVMFS~\cite{globalfs-cise-2015} filesystem is used to publish the software used by all of the major LHC
experiments on hundreds of computing sites around the world.
This requires that the computing site provide the CVMFS software,
and permit worker nodes to access the global network.

Figure~\ref{choices} outlines several broad methods for
accessing such global filesystems.  First, the most desirable
method is to {\bf mount} the filesystem directly from worker
nodes, so that the necessary files can be accessed on demand.
While this is possible at sites that are especially configured
to support the LHC, it is often not possible at other HPC sites
that restrict general network access from worker nodes.
There are many social and technical reasons for these restrictions, but they generally boil down to either security (avoiding intrusions into worker nodes) or efficiency (discouraging high value workers from waiting on network access.)  As a result, many HPC sites are unable
to use global software repositories.

If that cannot be accomplished, then the next
alternative is to {\bf copy} the entire repository to
a shared parallel or distributed filesystem at the site.
This has the advantage that the filesystem is already
mounted and available at worker nodes.  The entire
repository must be duplicated before it can be used,
regardless of whether an individual job needs it all.
A more subtle problem is that most parallel filesystems are
designed for delivering very large files, and are not at all 
efficient at delivering deep directory trees that
consist of small files and large amounts of metadata.
When a large number of workers begin to use the software
simultaneously, the system can collapse from the metadata
load.  ~\cite{metafs-pdsw-2017}  (XXX CITE MORE HERE)

A {\bf container} is often used to reduce the metadata load
on a parallel filesystem.  If a complex software tree is
deployed into a container, the underlying filesystem will
treat the container as a single large file, and be more
efficient at delivering it to many worker nodes simultaneously.
However, if the entire repository (measured in TB) is placed into
the container, it is likely to exceed a number of practical
limits on container size.  Individual worker nodes may have limited
local disk space, or may be entirely diskless.  Even if the
large container fits, it is likely that a given job does not need
\emph{all} of the repository simultaneously, and so it is wasteful
to transfer unneeded data.

To have containers of reasonable size, it is necessary to {\bf project}
the required subset of the global repository into a container of reasonable size that can be transferred to each worker node in an
efficient way. Of course, each job that depends upon a repository may need a different subset of the files, so it will be necessary to
maintain multiple container images, within the constraints of available 
storage space.

In this paper, we explore the technique of projecting a global
filesystem into local containers, in order to efficiently use
an HPC facility for a large number of jobs.  This requires striking
a balance between the size of containers, the amount of storage
space used, and the efficiency of data transfer.
We consider two phases of the problem: first, how to describe
and construct a container that serves the needs of a given job,
and second, how to efficiently manage resources when multiple
jobs and multiple containers are required.

@@

\if 0
\begin{table*}[t]
\begin{center}
\begin{tabular}{|c|r r r|r r r|}
\hline
CVMFS & & ls -lr & & & Shrinkwrap & \\
Repository & Total Size (MB) & inode Count & Time to walk & 
  Data Files & Deduplicated Data Files & Deduplicated Data (MB) \\ \hline
%alice-ocdb & 482,642,307,428 & 1,621,661 & 2:22 &
% 1,621,282 & 1,595 & 24,433,131,203\\
%alice & 3,980,875,747,986 & 27,322,350 & 51:53 & 
%  797,404 & 23,996,541 & 2,098,509,015,172\\
alice & 3,980,875 & 27,322,000 & 51:53 & 
  797,000 & 23,997,000 & 2,098,509\\
%atlas-condb & 719,548,941,240 & 11,781 & 0:04 &
%  10,940 & 229 & 471,492,086 \\
%atlas & 6,195,013,689,056 & 120,598,601 & 3:48:07 &
% 8,823,179 & 85,554,444 & 3,760,712,059,548 \\
atlas & 6,195,013 & 120,599,000 & 3:48:07 &
 8,823,000 & 85,554,000 & 3,760,712 \\
%boss & 120,048,527,315 & 2,418,025 & 4:18 &
% 256,061 & 1,868,192 & 72,301,955,666\\
boss & 120,048 & 2,418,000 & 4:18 &
 256,000 & 1,868,000 & 72,302\\
%cms & 581,789,406,847 & 26,224,567 & 1:07:30 &
% 18,667,178 & 101,448,867 & *2,242,658,825,701\\
cms & 581,789 & 26,225,000 & 1:07:30 &
 18,667,000 & 101,449,000 & *2,242,658\\
%geant4 & 297,571,283,648 & 7,128,024 & 10:36 &
% 323,280 & 6,219,891 & 68,381,969,660 \\
geant4 & 297,571 & 7,128,000 & 10:36 &
 323,000 & 6,220,000 & 68,382 \\
%grid & 30,642,826,259 & 714,235 & 1:17 &
% 141,622 & 583,923 & 18,804,003,040 \\
grid & 30,643 & 714,000 & 1:17 &
 142,000 & 584,000 & 18,804 \\
%lhcb & 2,134,893,079,440 & 33,736,802 & 1:01:26 &
% 10,032,339 & 20,406,235 & 368,070,277,634 \\
lhcb & 2,134,893 & 33,737,000 & 1:01:26 &
 10,032,000 & 20,406,000 & 368,070 \\
%na61 & 5,222,967,734 & 25,602 & 0:02 &
% 8,609 & 15,001 & 2,425,507,431 \\
na61 & 5,222 & 26,000 & 0:02 &
 9,000 & 15,000 & 2,426 \\
%sft & 3,996,527,275,237 & 128,608,792 & 4:17:58 &
% 6,821,699 & 63,210,139 & 2,016,609,810,573 \\ 
sft & 3,996,527 & 128,609,000 & 4:17:58 &
 6,822,000 & 63,210,000 & 2,016,610 \\ 
 \hline
\end{tabular}
\caption{CVMFS repository sizes.}
\label{tab:repo-sizes}
This table shows the size, number of inodes (rounded to nearest thousand), and the amount
of duplicate data that is eliminated using CVMFS.
\end{center}
\end{table*}

\begin{table}
\begin{center}
\begin{tabular}{|c|r r r|}
\hline
Repository & Total Size & inode Count & Deduplicated \\
 & (GB) & (K) & Data (GB) \\ \hline
alice & 3,981 & 27,322 & 2,099 \\
atlas & 6,195 & 120,599 & 3,761 \\
boss & 120 & 2,418 & 72\\
cms & 582 & 26,225 & *2,243\\
geant4 & 298 & 7,128 & 68 \\
grid & 31 & 714 & 19 \\
lhcb & 2,135 & 33,737 & 368,070 \\
na61 & 5 & 26 & 2 \\
sft & 3,997 & 128,609 & 2,017 \\ 
 \hline
\end{tabular}
\caption{CVMFS repository sizes.}
\label{tab:repo-sizes}
This table shows the size, number of inodes (rounded to nearest thousand), and the amount
of duplicate data that is eliminated using CVMFS.
\end{center}
\end{table}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{curated/lxplus_dist/usage_distribution_cdf.pdf}
    \caption{Usage distribution pulled from CERN LXPlus interactive machines.}
    \label{fig:usage_distribution}
\end{figure}
\if 0
spread across everyone, lot of commonality/overlap
few with very high usage, very large number of rarely used
lot not observed
\fi

@@

@@

\begin{figure*}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses.pdf}
\label{fig:sensitivity}
\if 0
[t] "just buy more cache"
efficiency is worse, since cache holds on to lots of questionable data
loads of opportunities to try merges, so we see a lot of unnecessary merges
less force removing bad choices, so more cruft can accumulate in many images
better cache utilization with tiny cache, but of course more thrashing
at scale, bad idea
large cache is too pricey at scale
[m] varying reuse doesn't affect too much
below 100 doesn't seem full/steady state yet
cache isn't settled? so haven't dropped bad merges; more unique data in cache boosts efficiency
[b] high number of requests, repeats don't matter
unique data dominates
hard to tell difference between [m] and [b], so not too sensitive to these
once you hit steady state, jobs count/repeats has little effect on efficiency
general trends of merging, not interesting by themselves
\fi
\end{figure*}

@@

\section{Related Work}
